Analyzing the Effect of Surface in Tennis Grand Slams

### 2018-09-08 contest/surface-type/paper-tennis.Rmd
Fixed games won and added data cleaning to package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)
```

```{r data-load, include = FALSE}
## Load WTA and ATP matches from deuce package
data(atp_matches)
data(wta_matches)
wta_matches$league <- "WTA"
atp_matches$league <- "ATP"

## Combining WTA and ATP
combined_tennis <- rbind(wta_matches, atp_matches)

## Subsetting from 2013 through 2017 inclusive, all the grand slams
gs <- combined_tennis %>% filter(tourney_level == "Grand Slams" & year >= 2013 & year <= 2017)
## Get rid of women's qualifiers
gs <- gs %>% filter(!(round %in% c("Q1", "Q2", "Q3", "Q4")))
dim(gs)

saveRDS(gs, "../../data/gs_2013_2017.rda")

@@

```{r data-cleaning, include = FALSE}
## Add new columns of winner points won, loser points won, and whether the winner had more points
gs <- gs %>% mutate(w_pointswon = w_1stWon + w_2ndWon + l_svpt - l_1stWon - l_2ndWon)
gs <- gs %>% mutate(l_pointswon = l_1stWon + l_2ndWon + w_svpt - w_1stWon - w_2ndWon)
gs <- gs %>% mutate(w_morepts = w_pointswon >= l_pointswon)

## Extract number of sets in each match
no_ret <- gsub(" RET", "", gs$score)
set_scores <- str_split(no_ret, " ")
n_sets <- sapply(set_scores, length)
gs$n_sets <- n_sets

## Data cleaning the tournament names
gs$tournament <- factor(gs$tourney_name)
gs$tournament <- gs$tournament %>% fct_collapse("US Open" = c("US Open", "Us Open"), 
                                                "French Open" = c("French Open", "Roland Garros"))
```

```{r games-won, include = FALSE, warning = FALSE}
## Extract games won for winner and loser
score_list <- strsplit(gs$score, " ")
games_won <- sapply(score_list, strsplit, split = "-|\\(")
gs$w_gameswon <- sapply(games_won, function(list){
  sum(as.integer(sapply(list, "[", 1)), na.rm = TRUE)
})
gs$l_gameswon <- sapply(games_won, function(list){
  sum(as.integer(sapply(list, "[", 2)), na.rm = TRUE)
})

```{r sets, include = FALSE, warning = FALSE}
## Extract number of sets for each match
no_ret <- gsub(" RET", "", gs$score)
set_scores <- str_split(no_ret, " ")
n_sets <- sapply(set_scores, length)
gs$n_sets <- n_sets

```{r bp-lost, include = FALSE}
gs$w_bplost <- gs$w_bpFaced - gs$w_bpSaved
gs$l_bplost <- gs$l_bpFaced - gs$l_bpSaved
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(deuce)
library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(kableExtra)

## To change into library(courtsports)
devtools::load_all("../..")
```

```{r data-cleaning, include = FALSE}
## Load WTA and ATP matches from deuce package
data(atp_matches)
data(wta_matches)

## Clean data
gs <- clean_matches_data(atp_matches = atp_matches,
                         wta_matches = wta_matches,
                         start_year = 2013,
                         end_year = 2017,
                         include_quals = FALSE,
                         level = "Grand Slams")

## Extract sets, points, and games
gs <- extract_sgp(gs)

# devtools::use_data(gs, overwrite = TRUE)

## Make into player data
gs_players <- matches_to_player_data(gs,
                                     w_opp_vars = "loser_rank",
                                     l_opp_vars = "winner_rank",
                                     opp_var_names = "opponent_rank")

# devtools::use_data(gs_players, overwrite = TRUE)

@@
